### 2024-10-09 the-good-the-bad-the-ugly.qmd

---
title: "IRT procedures for shortening tests"
author: "Ottavia M. Epifania & Friends"
format: 
  revealjs: 
    theme: mytheme.scss
background-opacity: "0.45"
title-slide-attributes: 
   data-background-image: "https://psicostat.github.io/shine-bright/gbu.jpg"
   data-background-size: 350px, cover
   data-background-position: 50% 5%, center
---

```{r, setup, include=FALSE}
library(knitr)

```

## Item Locating Algotihm -- ILA
### The bad

:::: {.columns}

::: {.column width="40%"}

<font size="4">
$B$: Item bank

$Q^k \subset B \setminus Q^{k-1}$: Set of item indexes selected for inclusion in the STF up to iteration $k$ ($Q^0 = \emptyset$)

$\mathbf{TIF}^*$: TIF target

$TIF^k = \frac{\sum_{i\in Q^k} IIF_i}{||Q^k||}$, where $||Q^k||$ denotes the cardinality of $Q^k$, $\mathbf{TIF}^0 = (0, 0, \ldots, 0)$

</font>

:::

::: {.column width="60%"}

```{r}
#| fig-align: center
knitr::include_graphics("img/ila.png")
```

:::

::::

## Frank[^1]
### The good

[^1]: **FRANK**ly I didnt' know how to name it

## Bruto 
### The bad

## Bullets

When you click the **Render** button a document will be generated that includes:

-   Content authored with markdown
-   Output from executable code

## Code

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```