### 2024-10-29 paper/algorithms.tex
update

\documentclass[12pt, a4paper, titilepage]{article}
\usepackage{graphicx}
\usepackage{subcaption}
\graphicspath{{C:/Users/huawei/Desktop/imag es/}} \usepackage{easyReview}
\usepackage{setspace}
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{apacite}
\usepackage{titlesec}
\usepackage{dcolumn}
\usepackage{lscape}
\usepackage{pdflscape}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{times}
\usepackage{booktabs}
\usepackage{fancyhdr}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{dcolumn}
\usepackage{siunitx}
\usepackage{array}
\usepackage{longtable}
\usepackage{lineno}
\usepackage{array}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{regexpatch}
\usepackage{multicol}
\usepackage{bm}
\usepackage{lineno}
\usepackage{color, colortbl} %righe colorate
\usepackage{lineno}
\usepackage{easyReview}
\usepackage{amssymb}

\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black]
\tikzstyle{decision} = [diamond, minimum width=1.5cm, minimum height=0.5cm, text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]

%opening
\title{The good, the bad, and the ugly}
\author{}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\newpage

\doublespacing

\section{Bruto}

%$ \theta = \{ x \in \mathbb{R} \,|\, \theta_{min} \leq x \leq \theta_{max} \}$: Latent trait levels, where $\theta_{min}$ is the minimum value of $\theta$ and $\theta_{max}$ is its maximum value \add{(non sono sicura della x)}

%$||\theta||$: cardinality of the latent trait

$||X||$: cardinality of $X$

$\forall Q \in\mathcal{Q} = 2^B \setminus \{\emptyset, B\}$,

\begin{enumerate}
	\item $\mathbf{TIF}^{Q} =  \frac{\sum_{i \in Q} IIF_i}{||Q||}$, where $||Q||$ is the cardinality of set $Q$
	\item $\overline{\Delta}_{\mathbf{TIF}^{Q}} =  \mathit{mean}(|\mathbf{TIF}^* - \mathbf{TIF}^{Q}|)$ %, that is the mean across all $\theta$ levels of the absolute difference between $TIF^*$ and $TIF^Q$ %\add{ho dei dubbi su questa notazione. la media è calcolata attraverso i valori di theta}. In alternativa a questa notazione: $\overline{\Delta}_{\mathbf{TIF}^{Q}} = \frac{|\mathbf{TIF}^* - \mathbf{TIF}^{Q}|}{||\theta||}$

\end{enumerate}

$Q_{bruto} = \arg \min_{Q \in \mathcal{Q}} \overline{\Delta}_{\mathbf{TIF}^{Q}}$

\add{spiegazione dell'algoritmo a parole}

\section{Item Locating Algorithm}

\add{Siccome la filosofia di ILA e ISA è molto simile sarebbe carino trovare loro un nome comune e poi declinarle nelle loro specificità. ci pensiamo}

\textbf{Set up: }

$B$: Item bank

$Q^k \subset B$: Set of item indexes selected for inclusion in the STF up to iteration $k$

$\mathbf{TIF}^*$: TIF target

$i^*$: Item selected at each iteration

%$\mathbf{pTIF}^k = \frac{\sum_{i\in Q^{k-1} \cup \{i^*\}} IIF_i}{||Q^{k-1}||+1}$, mean tif at iteration $k$, including the item selected in $D$

$||Q^k||$: cardinality of $Q^k$ at iteration $k$

At $k = 0$: $\mathbf{TIF}^0(\theta) = 0 \, \forall \theta$, $Q^0 = \emptyset$. For $k \geq 0$,

\begin{enumerate}
	\item $\theta_{target} := \arg \max |\mathbf{TIF}^* - \mathbf{TIF}^{k}|$
	\item $i^* = \arg \min_{i \in B\setminus Q^{k}} |\theta_{target} - b_i|$
	\item $\mathbf{pTIF}_{i^*} = \frac{TIF^k + IIF_{i^*}}{||Q^{k}|| + 1}$ %$\mathbf{pTIF}_{i^*}= \frac{\sum_{i \in Q^{k} \cup \{i^*\}} IIF_i}{||Q^{k} \cup \{i^*\}||}$ $=$ 
	\item Termination Criterion: $|\mathbf{TIF}^* - \mathbf{pTIF}_{i^*}| \geq |\mathbf{TIF}^* - \mathbf{TIF}^{k}|$: 
	\begin{itemize}
		\item FALSE:  $Q^{k+1} = Q^{k} \cup \{i^*\}$, $TIF^{k+1} = pTIF_{i^*}$, iterates 1-4 
		\item TRUE: Stop, %The item in $i^*$ does not contribute to reduce the distance from the TIF target, hence: 
		$Q_{ILA} = Q^k$
	\end{itemize}
\end{enumerate}

\section{Item Selecting Algorithm}

Same as ILA but based on the Item Information Functions.

\color{blue}

Set up same as ILA: 
$B$: Item bank

$Q^k \subset B$: Set of item indexes selected for inclusion in the STF up to iteration $k$

$\mathbf{TIF}^*$: TIF target

$i^*$: Item selected at each iteration

%$\mathbf{pTIF}^k = \frac{\sum_{i\in Q^{k-1} \cup \{i^*\}} IIF_i}{||Q^{k-1}||+1}$, mean tif at iteration $k$, including the item selected in $D$

$||Q^k||$: cardinality of $Q^k$ at iteration $k$

\normalcolor

At $k = 0$: $\mathbf{TIF}^0(\theta) = 0 \, \forall \theta$, $Q^0 = \emptyset$. For $k \geq 0$,

\begin{enumerate}
	\item $\theta_{target} := \arg \max |\mathbf{TIF}^* - \mathbf{PIF}_D^{k}|$
%	\item $IIF_{i \in B \setminus Q^k}(\theta_{target}) = a_i^2P(\theta_{target}, a_i, b_i)[1-P(\theta_{target}, a_i, b_i)]$
	\item $i^* := \arg \max_{i \in B\setminus Q^k} IIF_i(\theta_{target})$
	\item $\mathbf{pTIF}_{i^*} = \frac{TIF^k + IIF_{i^*}}{||Q^{k}|| + 1}$
	\item Termination Criterion: $|\mathbf{TIF}^* - \mathbf{PIF}_D^k| \geq |\mathbf{TIF}^* - \mathbf{TIF}^{k}|$: 
	\begin{itemize}
		\item TRUE: $k := k +1$, $Q^{k} = Q^{k-1} \cup \{D\}$, iterates 1-5 (\add{Ho scritto $k-1$ perché siccome ho scritto all'inizio che $k$ si aggiorna e diventa $k+1$ mi sembrava sensato})
		\item FALSE: The item in $D$ does not contribute to reduce the distance from the TIF target, hence: $Q_{ISA} = Q^k$
	\end{itemize}
\end{enumerate}

\section{Frank}

The setup is like the one of ILA and ISA:

$B$: Item bank

$Q^k$: set of items selected at iteration $k$

$i^*$: provisional item selected at each iteration

\add{$\mathbf{PIF}$: provisional mean tif}

At $k =0$, $\mathbf{PIF}^0 = (0, 0, \ldots, 0)$, $Q^0 = \emptyset$, iterate
\alert{gira il cirterio di terminazione e uniforma con gli altri}
\begin{enumerate}
	\item  $A^k = B \setminus Q^k$ (sets of available items at iteration $k$)
	\item $\forall i \in A^k$, $\mathbf{PIF}_{i}^k = \frac{\mathbf{TIF}^k + \mathbf{IIF}_{i}}{||Q^k||+1}$
	\item $D = \arg \min_{i \in A^k} |\mathbf{TIF}^* - \mathbf{PIF}_i|$
	\item Termination criterion: $|\mathbf{TIF}^* - \mathbf{PIF}_D^{k}| < |\mathbf{TIF}^* - \mathbf{TIF}^{k}|$: 
	\begin{itemize}
		\item  If true, $k := k + 1,$ $Q^{k} = Q^{k-1} \cup \{D\}$, restart from 1 \add{C'è il meno 1 per la stessa ragione scritta per ILA e ISA}

		\item  If false, stops, $Q_{Frank} = Q^k$

	\end{itemize}
\end{enumerate}

\end{document}